---
title: "Tom Bliss Tutorial"
author: "Nate Rowan"
date: "10/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is adapted from Tom Bliss's tutorial on Kaggle.

First, we load in the required libraries.

```{r, warning=FALSE}
library(tidyverse)
library(gganimate)
library(cowplot)
library(repr)
```

Then, we read in the data.

```{r}
##reading in non-tracking data

#includes schedule info for games
df_games <- read_csv(here::here("datasets/games.csv"),
                     col_types = cols())

#includes play-by-play info on specific plays
df_plays <- read_csv(here::here("datasets/plays.csv"),
                     col_types = cols())

#includes background info for players
df_players <- read_csv(here::here("datasets/players.csv"),
                       col_types = cols())


##Reading tracking data (needs to be done iteratively)

df_tracking <- read_csv(here::here("datasets/week1.csv"),
                        col_types = cols())
```

Next, we standardize field coordinates to ensure that the offense is always going in the same direction.

```{r}
### Standardize tracking data so it's always in direction of offense
### vs raw on-field coordinates
df_tracking <- df_tracking %>%
  mutate(x = ifelse(playDirection == "left", 120-x, x),
         y = ifelse(playDirection == "left", 160/3 - y, y))
```

Then, we are going to animate a couple plays. First, we declare our field boundaries that will be used in our plots.

```{r}
### Declare values for general field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3
```

Next, we select a play and merge it with our games data and tracking data

```{r}
#picking a random play
set.seed(999)

### Select games from week 1, since I only loaded in week 1 tracking data
df_games_week1 <- df_games %>%
  filter(week == 1)

### Only select plays that occurred in week 1
df_plays_week1 <- semi_join(df_plays, df_games_week1, by = "gameId")

### Select a random play from week 1
example_play <- df_plays_week1 %>%
                select(gameId, playId, playDescription) %>% 
                sample_n(1)


### merging games data to play
example_play <- inner_join(example_play,
                          df_games,
                          by = c("gameId" = "gameId"))


### merging tracking data to play
example_play <- inner_join(example_play,
                           df_tracking,
                           by = c("gameId" = "gameId",
                                  "playId" = "playId"))

```

Now I have the play, game, and tracking data from this play in week 1 of 2018. 

Now, we can create our plot of the play.

```{r}
### colors for plot. Order: offense, football, defense
cols_fill <- c("#003594", "#663300", "#A5ACAF")
cols_col <- c("#000000", "#663300", "#000000")

### play description will be title of plot
### gsub() replaces a regex with something, so this seems to
### be replacing anything with parenthesis with nothing
plot_title <- str_trim(gsub("\\s*\\([^\\)]+\\)","",as.character(example_play$playDescription[1])))

### Specific boundaries for a given play
ymin <- max(round(min(example_play$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(example_play$x, na.rm = TRUE) + 10, -1), 120)

### Hash Marks
df.hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df.hash <- df.hash %>% filter(!(floor(y %% 5) == 0))
df.hash <- df.hash %>% filter(y < ymax, y > ymin)


### Time to actually plot!
ggplot() +

#setting size and color parameters
scale_size_manual(values = c(6, 4, 6), guide = FALSE) + 
scale_shape_manual(values = c(21, 16, 21), guide = FALSE) +
scale_fill_manual(values = cols_fill, guide = FALSE) + 
scale_colour_manual(values = cols_col, guide = FALSE) +

#adding hash marks
annotate("text", x = df.hash$x[df.hash$x < 55/2], 
       y = df.hash$y[df.hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
annotate("text", x = df.hash$x[df.hash$x > 55/2], 
       y = df.hash$y[df.hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 

#adding yard lines
annotate("segment", x = xmin, 
       y = seq(max(10, ymin), min(ymax, 110), by = 5), 
       xend =  xmax, 
       yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 

#adding field yardline text
annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
       label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
       angle = 270, size = 4) + 
annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
       label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
       angle = 90, size = 4) + 
  
#adding field exterior
annotate("segment", x = c(xmin, xmin, xmax, xmax), 
       y = c(ymin, ymax, ymax, ymin), 
       xend = c(xmin, xmax, xmax, xmin), 
       yend = c(ymax, ymax, ymin, ymin), colour = "black") + 

#adding players
geom_point(data = example_play, aes(x = (xmax-y),
                                  y = x, 
                                  shape = team,
                                  fill = team,
                                  group = nflId,
                                  size = team,
                                  colour = team), 
         alpha = 0.7) +  

#adding jersey numbers
geom_text(data = example_play, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
        vjust = 0.36, size = 3.5) + 

#applying plot limits
ylim(ymin, ymax) + 
coord_fixed() +
  
  #applying theme
theme_nothing() + 
theme(plot.title = element_text()) +

#titling plot with play description
labs(title = plot_title) +

#setting animation parameters
transition_time(frameId)  +
ease_aes('linear') + 
NULL
```



